Towards Performance-portable Parallel Programming in Computer Science Education

### 2020-01-21 src/2.Kokkos.tex
UVM memory -> unified memory

To capture semantic information, that is the~\emph{What}, the ~\emph{Where} and the ~\emph{How}, Kokkos introduces six abstractions: \emph{execution spaces}, \emph{execution patterns}, \emph{execution policies}, \emph{memory spaces}, \emph{memory layout} and \emph{memory traits}. These abstractions specify semantic information that enable to capture the programmer's intent and the runtime to efficiently map the program to any underlying hardware architecture. We list them as follows:
\begin{itemize}
	\item  An execution space is a place where code can be executed. On current hardware architectures this correspond to accelerators and CPUs and can include any compute device in the future. This abstraction supports remote compute devices in distributed memory scenarios as remote execution spaces.
	\item Execution patterns expose the parallel programming paradigm. Supported patterns are \emph{parallel\_for} loop that executes the loop body in any order a specified amount of times, the \emph{parallel\_reduce} which combines a parallel\_for with a reduction operation, \emph{parallel\_scan} which combines a parallel\_for operation with a prefix or postfix scan, and \emph{task} which executes a single function potentially in parallel in respect to other tasks or code sections. 
	\item Execution policy shape the iteration space of a loop pattern. A simple execution policy is a range policy. It specifies that the loop body is executed once for each element in a range.
	\item Memory spaces are the places where data resides. They specify physical locations of data as well as access characteristics. Different physical locations correspond to different device types such as high bandwidth memories, on-die scratch memories or non-volatile bulk storage. Different logical memory spaces allow for concepts such as UVM memory in the CUDA programming model, which is accessible from the host and the CUDA accelerator. Memory spaces, similarly to execution spaces, conceptually support remote memory locations in distributed-memory scenarios. Furthermore, they encapsulate functionality such as consistency control and persistence scopes.
	\item Layouts express the mapping from loop indices to address offsets. By adopting appropriate layouts for memory structures, an application can optimize data access patterns in a given algorithm. If an implementation provides polymorphic layouts (i.e. a data structure can be instantiated at compile or runtime with different layouts), architecture-dependent optimizations can be performed.
	\item Memory traits specify how a data structure is accessed. Traits express usage scenarios such as atomic access, random access and streaming loads or stores. This allows the programming model to  optimize load and store operations.
\end{itemize}

>>>>

To capture semantic information, that is the~\emph{What}, the ~\emph{Where} and the ~\emph{How}, Kokkos introduces six abstractions: \emph{execution spaces}, \emph{execution patterns}, \emph{execution policies}, \emph{memory spaces}, \emph{memory layout} and \emph{memory traits}. These abstractions specify semantic information that enable to capture the programmer's intent and the runtime to efficiently map the program to any underlying hardware architecture. We list them as follows:
\begin{itemize}
	\item  An execution space is a place where code can be executed. On current hardware architectures this correspond to accelerators and CPUs and can include any compute device in the future. This abstraction supports remote compute devices in distributed memory scenarios as remote execution spaces.
	\item Execution patterns expose the parallel programming paradigm. Supported patterns are \emph{parallel\_for} loop that executes the loop body in any order a specified amount of times, the \emph{parallel\_reduce} which combines a parallel\_for with a reduction operation, \emph{parallel\_scan} which combines a parallel\_for operation with a prefix or postfix scan, and \emph{task} which executes a single function potentially in parallel in respect to other tasks or code sections. 
	\item Execution policy shape the iteration space of a loop pattern. A simple execution policy is a range policy. It specifies that the loop body is executed once for each element in a range.
	\item Memory spaces are the places where data resides. They specify physical locations of data as well as access characteristics. Different physical locations correspond to different device types such as high bandwidth memories, on-die scratch memories or non-volatile bulk storage. Different logical memory spaces allow for concepts such as memory in the CUDA programming model, which is accessible from the host and the CUDA accelerator. Memory spaces, similarly to execution spaces, conceptually support remote memory locations in distributed-memory scenarios. Furthermore, they encapsulate functionality such as consistency control and persistence scopes.
	\item Layouts express the mapping from loop indices to address offsets. By adopting appropriate layouts for memory structures, an application can optimize data access patterns in a given algorithm. If an implementation provides polymorphic layouts (i.e. a data structure can be instantiated at compile or runtime with different layouts), architecture-dependent optimizations can be performed.
	\item Memory traits specify how a data structure is accessed. Traits express usage scenarios such as atomic access, random access and streaming loads or stores. This allows the programming model to  optimize load and store operations.
\end{itemize}
